---
title: "spotify"
author: "Thom Tran"
date: "2022-09-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
### About business
Spotify transformed music listening forever when it launched in 2008. Discover, manage and share over 80 million tracks, including more than 4 million podcast titles, for free, or upgrade to Spotify Premium to access exclusive features for music including improved sound quality and an on-demand, offline, and ad-free music listening experience. Today, Spotify is the world’s most popular audio streaming subscription service with 433m users, including 188m subscribers, across 183 markets.

### Tasks

## Prepare
### Data source

The data set used in this analyze is [Spotify Data Visulization](https://www.kaggle.com/code/varunsaikanuri/spotify-data-visualization) on Kaggle

## Process

First we load necessary package

```{r}
install.packages("tidyverse")
install.packages("lubridate")
install.packages("dplyr")
install.packages("corrplot")
library(tidyverse)
library(lubridate)
library(dplyr)
library(corrplot)

```
set working directory
```{r}
setwd("/Users/HM3/Downloads")
```

load data set
```{r}
song <- read.csv("songs_normalize.csv")
```

Next, we will check column name in data set
```{r}
colnames(song)
```
Explain variables in data set
* artist: Name of the Artist.
* song: Name of the Track.
* duration_ms: Duration of the track in milliseconds.
* explicit: The lyrics or content of a song or a music video contain one or more of the criteria which could be considered offensive or unsuitable for children.
* year: Release Year of the track.
* popularity: The higher the value the more popular the song is.
* danceability: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable.
* energy: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
* key: The key the track is in. Integers map to pitches using standard Pitch Class notation. E.g. 0 = C, 1 = C♯/D♭, 2 = D, and so on. If no key was detected, the value is -1.
* loudness: The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude). Values typically range between -60 and 0 db.
* mode: Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.
* speechiness: Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. Values above 0.66 describe tracks that are probably made entirely of spoken words. Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music. Values below 0.33 most likely represent music and other non-speech-like tracks.
* acousticness: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
* instrumentalness: Predicts whether a track contains no vocals. "Ooh" and "aah" sounds are treated as instrumental in this context. Rap or spoken word tracks are clearly "vocal". The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content. Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0.
* liveness: Detects the presence of an audience in the recording. Higher liveness values represent an increased probability that the track was performed live. A value above 0.8 provides strong likelihood that the track is live.
* valence: A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).
* tempo: The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.
* genre: Genre of the track.

Glimspe at the data set
```{r}
glimpse(song)
```
Check null values in columns
```{r}
sapply(song, function(x) sum(is.na(x)))
```

there are no missing values in dataframe

check row dupicate in song column
```{r}
which(duplicated(song$song))
```

delete duplicate
```{r}
new_song <- song[!duplicated(song$song), ]
```

check duplicate again in new_song
```{r}
which(duplicated((new_song)))
```
Summarize the new_song data set
```{r}
str(new_song)
```


## Analyze

First, i will check correlation between variables
```{r}
new_song1 <- new_song[ ,c(3,5,6,7,8,9,10,11,12,13,14,15,16,17)]
library(corrplot)
corrplot(cor(new_song1), method = 'square', title = "correlation of columns", tl.cex = 0.75)


```
valence ~ danceability, loudness ~ energy, valence ~ energy, valence ~ loudness

Create histogram chart for each column to find 
```{r}
hist(new_song$duration_ms, main = 'Histogram of duration', xlab = 'Duration')
hist(new_song$popularity, main = 'Histogram of popularity', xlab = 'popularity')
hist(new_song$danceability, main =  'Histogram of danceability', xlab = 'danceability')
hist(new_song$energy, main = 'Histogram of energy', xlab = 'energy')
hist(new_song$loudness, main = 'Histogram of loudness', xlab = 'loudness')
hist(new_song$key, main = 'Histogram of key', xlab = 'key')
hist(new_song$speechiness, main = 'Histogram of speechiness', xlab = 'speechiness')
hist(new_song$acousticness, main = 'Histogram of acousticness', xlab = 'acousticness')
hist(new_song$liveness, main = 'Histogram of liveness', xlab = 'liveness')
hist(new_song$valence, main = 'Histogram of valence', xlab = 'valence')
hist(new_song$tempo, main = 'Histogram of tempo', xlab = 'tempo')
```
Relation between loudness and valence
```{r}
ggplot(new_song, aes(valence, loudness)) + geom_point(color = 'darkred') + geom_smooth(method = 'lm')
```
we see that loudness have no relation with valence
Next, we will find relationship between tempo and dancebility
```{r}
ggplot(new_song, aes(tempo, danceability)) +geom_point(color = 'darkblue') + geom_smooth(method = 'lm')
```
negative relationship => speed of the beat more high, the danceability more low
Relation between energy and loudness
```{r}
ggplot(new_song, aes(energy, loudness)) +geom_point(color = 'bisque3') + geom_smooth(method = 'lm')
```
energy have positive relation with loudness
relation between acoustic and danceability
TOP 10 ARTIST NAME 
```{r}
top_artist <- new_song %>%
  count(artist) %>%
  group_by(artist) %>%
  arrange(desc(n)) %>%
  head(10)
head(top_artist, 10)
ggplot(top_artist, aes(x= reorder(artist, -n), y = n)) + geom_bar(stat = 'identity', fill = 'darkblue')+ labs(title = 'TOP 10 ARTIST', x = 'Name of artist', y = 'Number of song') + theme(axis.text.x = element_text(angle = 30), plot.title =  element_text(hjust = 0.5))
```
TOP 10 TRACK
```{r}
top_track <- new_song %>%
  select(song, popularity) %>%
  arrange(-popularity) %>%
  head(10)
head(top_track, 10)
ggplot(top_track, aes(x = reorder(song, -popularity), y = popularity, fill = song)) + geom_bar(stat = 'identity') + labs(title = 'TOP 10 TRACK', x = 'Name of Tracks', y = 'Score of popularity') + theme(axis.text.x = element_text(angle = 90), plot.title = element_text(hjust = 0.5))
```

EFFECT OF DACNCEBILITY ON POPULARITY
```{r}
ggplot(new_song, aes(danceability, popularity )) + geom_point(color = 'red') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF DANCEABILITY ON POPULARITY")
```
We can see that danceability have no relation with popularity

EFFECT OF VALENCE ON POPULARITY
```{r}
ggplot(new_song, aes(valence, popularity)) + geom_point(color = 'darkgreen') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF VALENCE ON POPULARITY" )
```
We can see that valence have no relation with popularity
EFFECT OF ENERGY ON POPULARITY
```{r}
ggplot(new_song, aes(energy, popularity)) + geom_point(color = 'darkorange4') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF ENERGY ON POPULARITY" )
```
```{r}
ggplot(new_song, aes(loudness, popularity)) + geom_point(color = 'darkslateblue') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF LOUDNESS ON POPULARITY" )
```

```{r}
ggplot(new_song, aes(speechiness, popularity)) + geom_point(color = 'forestgreen') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF SPEECHINESS ON POPULARITY" )
```

```{r}
ggplot(new_song, aes(acousticness, popularity)) + geom_point(color = 'goldenrod4') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF ACOUSTICNESS ON POPULARITY" )
```
```{r}
ggplot(new_song, aes(tempo, popularity)) + geom_point(color = 'deepskyblue4') + geom_smooth(method = 'lm') + labs(title = "EFFECT OF TEMPO ON POPULARITY" )
```
```{r}
unique(new_song$genre)
```

